Parsing Postfix log files

### 2009-04-07 doc/logparser-implementation.tex
Consistent float placement.
Perfect oracle section is more or less finished, at last. Am now wondering if I should remove the perfect worst stuff.

\parsername{} deals with all the complications of parsing, the
eccentricities and oddities of Postfix log files, presenting the resulting
data in a normalised, simple to use representation.  The parser's task is
to follow the journey each mail takes through Postfix, combining the data
captured by rules into a coherent whole, saving it in a useful and
consistent form, and performing housekeeping duties.  The intermingling of
log entries from different mails immediately rules out the possibility of
handling each mail in isolation; the parser must be able to handle multiple
mails in parallel, each potentially at a different stage in its journey,
without any interference between mails --- except in the minority of cases
where intra-mail interference is required, e.g.\ mail re-injected for
forwarding (\sectionref{Re-injected mails}).  The best way to deal with
intermingling is to maintain state information for every unfinished mail
and manipulate the appropriate mail correctly for each log line
encountered.

@@

    \item [PICKUP] The \action{PICKUP} action corresponds to the
        \daemon{pickup} service dealing with a locally submitted mail.  A
        new mail will be created, although out of order log entries may
        have caused it to already exist, as documented in
        \sectionref{pickup logging after cleanup}.

@@

% do not reformat this!
\begin{table}[ht]
    \caption{Example of a rule used by \parsernameshort{}}
    \empty{}\label{Example rule in implementation table}
    \begin{tabular}{ll}
        \tabletopline{}%
        \textbf{Field}      & \textbf{Value}                                    \\
        \tablemiddleline{}%
        name                & Unknown sender domain                             \\
        description         & We do not accept mail from unknown domains        \\
        restriction\_name   & reject\_unknown\_sender\_domain                   \\
        program             & \daemon{smtpd}                                    \\
        regex               & \verb!^__RESTRICTION_START__ <(__SENDER__)>: !    \\
                            & \verb!Sender address rejected: Domain not found;! \\
                            & \verb!from=<\k<sender>> to=<(__RECIPIENT__)> !    \\
                            & \verb!proto=E?SMTP helo=<(__HELO__)>$!            \\
        result\_data        &                                                   \\
        connection\_data    &                                                   \\
        action              & \action{DELIVERY\_REJECTED}                       \\
        hits                & 0                                                 \\
        hits\_total         & 0                                                 \\
        priority            & 0                                                 \\
        debug               & 0                                                 \\
        %cluster\_group      & 400                                               \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

>>>>

\parsername{} deals with all the complications of parsing, the
eccentricities and oddities of Postfix log files, presenting the resulting
data in a normalised, simple to use representation.  The parser's task is
to follow the journey each mail takes through Postfix, combining the data
captured by rules into a coherent whole, saving it in a useful and
consistent form, and performing housekeeping duties.  The intermingling of
log lines from different mails immediately rules out the possibility of
handling each mail in isolation; the parser must be able to handle multiple
mails in parallel, each potentially at a different stage in its journey,
without any interference between mails --- except in the minority of cases
where intra-mail interference is required, e.g.\ mail re-injected for
forwarding (\sectionref{Re-injected mails}).  The best way to deal with
intermingling is to maintain state information for every unfinished mail
and manipulate the appropriate mail correctly for each log line
encountered.

@@

    \item [PICKUP] The \action{PICKUP} action corresponds to the
        \daemon{pickup} service dealing with a locally submitted mail.  A
        new mail will be created, although out of order log lines may have
        caused it to already exist, as documented in \sectionref{pickup
        logging after cleanup}.

@@

% do not reformat this!
\begin{table}[thbp]
    \caption{Example of a rule used by \parsernameshort{}}
    \empty{}\label{Example rule in implementation table}
    \begin{tabular}{ll}
        \tabletopline{}%
        \textbf{Field}      & \textbf{Value}                                    \\
        \tablemiddleline{}%
        name                & Unknown sender domain                             \\
        description         & We do not accept mail from unknown domains        \\
        restriction\_name   & reject\_unknown\_sender\_domain                   \\
        program             & \daemon{smtpd}                                    \\
        regex               & \verb!^__RESTRICTION_START__ <(__SENDER__)>: !    \\
                            & \verb!Sender address rejected: Domain not found;! \\
                            & \verb!from=<\k<sender>> to=<(__RECIPIENT__)> !    \\
                            & \verb!proto=E?SMTP helo=<(__HELO__)>$!            \\
        result\_data        &                                                   \\
        connection\_data    &                                                   \\
        action              & \action{DELIVERY\_REJECTED}                       \\
        hits                & 0                                                 \\
        hits\_total         & 0                                                 \\
        priority            & 0                                                 \\
        debug               & 0                                                 \\
        %cluster\_group      & 400                                               \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

### 2009-04-07 doc/logparser-results.tex
Perfect oracle section is more or less finished, at last. Am now wondering if I should remove the perfect worst stuff.

@@

XXX USE THE DATA COLLECTED ABOUT NUMBER OF RULES TRIED\@.  MAYBE INCLUDE IN
ORDERINGS SECTION\@?  MIX WITH RECOGNISING VS RECOGNISING + ACTION, AND
KEEP COUNTERS FOR EACH PROGRAM TOO --- MAYBE SOMETHING USEFUL WILL COME
FROM IT\@.  COMPARE NUMBER OF RULES TRIED FOR DIFFERENT RULE ORDERINGS\@.

XXX SHOULD I USE $\backslash{}$clearpage{} OR SOMETHING TO AFFECT FLOAT
PLACEMENT\@?  GRAPHS AND TABLES TEND TO MOVE QUITE FAR IN SOME OF THE
SUBSECTIONS, THOUGH MAKING THE TEXT LONGER MIGHT HELP TOO\@.  IDEALLY I
WANT TO HAVE TEXT, THEN ALL GRAPHS AND TABLES, THEN THE NEXT SECTION\@.

\newpage{} % XXX VARIOREF PROBLEMS

@@

\begin{table}[thbp]
    \caption{Details of the computer used to generate statistics}
    \empty{}\label{Details of the computer used to generate statistics}
    \begin{tabular}[]{ll}
        \tabletopline{}%
        Component  & Component in use                                   \\
        \tablemiddleline{}%
        CPU         & One dual core 2.40GHz Intel\textregistered{}
                        Core\texttrademark{}2 CPU,                      \\
                    & with 32KB L1 cache and 4MB L2 cache.              \\
        RAM         & 2GB 667 MHz DDR RAM\@.                            \\
        Hard disk   & One Seagate Barracuda 7200 RPM 250GB SATA disk.   \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

@@

An important property of a parser is how parsing time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
\Graphref{parsing time vs file size vs number of log lines graph} shows the
parsing time in seconds, file size in MB, and number of log lines in tens
of thousands, for each of the \numberOFlogFILES{} log files.  The three
lines run roughly in parallel, giving the impression that the algorithm
scales linearly with input size.  This impression is borne out by
\graphref{parsing time vs file size vs number of log lines factor}, which
plots both the ratio of file size vs parsing time, and the ratio of number
of log lines vs parsing time (higher is better in both cases);
\tableref{parsing time vs file size vs number of log lines factor table}
shows the same ratios for different groups of log files.  The ratios are
quite tightly banded, showing that the algorithm scales linearly; the
ratios increase (i.e.\ improve) for log files 22 and 62--68, despite their
larger than average size.  Both groups of log files are much larger than
usual due to a mail loop caused by a user who set up mail forwarding
incorrectly, resulting in a very different distribution of log lines:
normally most log lines are generated by mail delivery attempts from other
hosts, but when the mail loops occurred most of the log lines resulted from
failed delivery of mail generated on the server itself.  The Postfix
components that generated most of the log lines during the mail loop have
fewer associated rules than the Postfix components whose log lines normally
make up the bulk of each log file, so the average number of rules tried per
log line reduces and so does the average parsing time per log line.
\Graphref{Number of rules per Postfix component} shows the number of rules
per Postfix component, \graphref{Average number of rules tried per log
line} shows the drop in the average number of rules tried per log line for
log files containing a mail loop, and \graphref{Average number of rules
tried per log line for each Postfix component} shows the average number of
rules tried per log line for each Postfix component.

@@

\Graphref{Parsing time relative to shuffled ordering graph} shows the
parsing times of optimal and reverse orderings relative to shuffled
ordering; the mean relative parsing times for different groupings of log
files are given in \tableref{Parsing time relative to shuffled ordering
table}.  This optimisation provides a mean reduction in parsing time of
\input{build/include-optimal-ordering-parsing-time-reduction-other-logs}\%
with normal log files,
\input{build/include-optimal-ordering-parsing-time-reduction-logs-22-62-68}\%
when a mail loop occurs and the distribution of log lines is unusual.
Optimal rule ordering has other benefits, described in
\sectionref{scalability as the number of rules rises}.  \Tableref{Parsing
time relative to shuffled ordering table} shows that differences in rule
ordering have less effect on parsing time when parsing log files 22 and
62--68, due to the different distribution of log lines in those log files.

@@

\subsection{Perfect Rule Ordering}

@@

XXX WRITE ABOUT PERFECT RULE ORDERING\@.

@@

\Graphref{Parsing time relative to shuffled ordering including perfect best
and worst graph} shows the parsing times of the perfect best oracle and
perfect worst oracle relative to shuffled ordering, in addition to the
parsing times of optimal and reverse orderings.  As expected, the perfect
best oracle is more efficient than optimal ordering, and likewise the
perfect worst oracle is less efficient than reverse ordering.

XXX REFER TO SECOND AND THIRD GRAPHS, AND ADD TABLES SHOWING MEAN AND
STDDEV FOR THE THREE GROUPS OF LOGS FILES\@.

XXX WRITE SOMETHING TO MAKE IT EASIER TO GENERATE MEAN AND STDDEV FOR THREE
GROUPS OF LOG FILES\@.

\Graphref{Parsing time of perfect best oracle relative to optimal ordering}

\showgraph{build/graph-perfect-best-and-perfect-worst-and-optimal-and-reverse-vs-shuffled}{Parsing
times relative to shuffled ordering}{Parsing time relative to shuffled
ordering including perfect best and worst graph}

\showgraph{build/graph-perfect-best-vs-optimal}{Parsing time of perfect
best oracle relative to optimal ordering}{Parsing time of perfect best
oracle relative to optimal ordering}

\showgraph{build/graph-perfect-worst-vs-reverse}{Parsing time of perfect
worst oracle relative to reverse ordering}{Parsing time of perfect worst
oracle relative to reverse ordering}

\showtable{build/include-perfect-best-and-optimal-and-reverse-and-perfect-worst-vs-shuffle}{Parsing
time including perfect best and worst oracles relative to shuffled
ordering}{Parsing time including perfect best and worst oracles relative to
shuffled ordering table}

@@

How any architecture scales as the number of rules increases is important,
but it is particularly important in this architecture because it is
expected that the typical parser will have a large number of rules.  There
are \numberOFrules{} rules in the full \parsername{} ruleset, whereas the minimum
number of rules required to parse the \numberOFlogFILES{} log files used
when generating the results in this thesis is \numberOFrulesMINIMUM{},
\numberOFrulesMINIMUMpercentage{} of the full ruleset.  The full ruleset is
larger because \parsername{} is tested with \numberOFlogFILESall{} log
files, to improve its parsing and ensure it correctly parses logs generated
by later Postfix versions.  A second set of test runs was performed using
the minimum ruleset, and the parsing times compared to those generated
using the full ruleset: the percentage parsing time increase when using the
full ruleset instead of the minimal ruleset for optimal, shuffled and
reverse orderings is shown in \graphref{Percentage parsing time increase of
maximum ruleset over minimum ruleset}.

>>>>

% Reduce the gap between columns in tables.
\addtolength{\tabcolsep}{-1pt}

@@

@@

\begin{table}[thbp]
    \caption{Details of the computer used to generate statistics}
    \empty{}\label{Details of the computer used to generate statistics}
    \centering{}
    \begin{tabular}[]{ll}
        \tabletopline{}%
        Component  & Component in use                                   \\
        \tablemiddleline{}%
        CPU         & One dual core 2.40GHz Intel\textregistered{}
                        Core\texttrademark{}2 CPU,                      \\
                    & with 32KB L1 cache and 4MB L2 cache.              \\
        RAM         & 2GB 667 MHz DDR RAM\@.                            \\
        Hard disk   & One Seagate Barracuda 7200 RPM 250GB SATA disk.   \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

@@

An important property of a parser is how parsing time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
\Graphref{parsing time vs file size vs number of log lines graph} shows the
parsing time in seconds, file size in MB, and number of log lines in tens
of thousands, for each of the \numberOFlogFILES{} log files.  The three
lines run roughly in parallel, giving the impression that the algorithm
scales linearly with input size.  This impression is borne out by
\graphref{parsing time vs file size vs number of log lines factor}, which
plots both the ratio of file size vs parsing time, and the ratio of number
of log lines vs parsing time (higher is better in both cases);
\tableref{parsing time vs file size vs number of log lines factor table}
shows the same ratios for different groups of log files.  The ratios are
quite tightly banded, showing that the algorithm scales linearly; the
ratios increase (i.e.\ improve) for log files 22 \& 62--68, despite their
larger than average size.  Both groups of log files are much larger than
usual due to a mail loop caused by a user who set up mail forwarding
incorrectly, resulting in a very different distribution of log lines:
normally most log lines are generated by mail delivery attempts from other
hosts, but when the mail loops occurred most of the log lines resulted from
failed delivery of mail generated on the server itself.  The Postfix
components that generated most of the log lines during the mail loop have
fewer associated rules than the Postfix components whose log lines normally
make up the bulk of each log file, so the average number of rules tried per
log line reduces and so does the average parsing time per log line.
\Graphref{Number of rules per Postfix component} shows the number of rules
per Postfix component, \graphref{Average number of rules tried per log
line} shows the drop in the average number of rules tried per log line for
log files containing a mail loop, and \graphref{Average number of rules
tried per log line for each Postfix component} shows the average number of
rules tried per log line for each Postfix component.

@@

\Graphref{Parsing time relative to shuffled ordering graph} shows the
parsing times of optimal and reverse orderings relative to shuffled
ordering; the mean relative parsing times for different groupings of log
files are given in \tableref{Parsing time relative to shuffled ordering
table}.  This optimisation provides a mean reduction in parsing time of
\input{build/include-optimal-ordering-parsing-time-reduction-other-logs}\%
with normal log files,
\input{build/include-optimal-ordering-parsing-time-reduction-logs-22-62-68}\%
when a mail loop occurs and the distribution of log lines is unusual.
Optimal rule ordering has other benefits, described in
\sectionref{scalability as the number of rules rises}.  \Tableref{Parsing
time relative to shuffled ordering table} shows that differences in rule
ordering have less effect on parsing time when parsing log files 22 \&
62--68, due to the different distribution of log lines in those log files.

@@

\subsection{Comparing Optimal Ordering Against A Perfect Oracle}

@@

XXX SHOULD I BOTHER WITH PERFECT WORST\@?

@@

\Graphref{Parsing time relative to shuffled ordering including perfect best
and worst graph} shows the parsing times of the perfect best and perfect
worst oracles, and the optimal and reverse orderings relative to shuffled
ordering; \tableref{Parsing time relative to shuffled ordering, including
oracles table} shows mean and standard deviation.  As expected, the perfect
best oracle is more efficient than optimal ordering, and likewise the
perfect worst oracle is less efficient than reverse ordering.
\Graphref{Percentage increase in parsing time when using optimal ordering
instead of perfect best oracle graph} and \graphref{Percentage decrease in
parsing time when using reverse ordering instead of perfect worst oracle
graph} show the percentage increase and decrease in parsing time when
using, respectively, optimal ordering instead of the perfect best oracle,
and reverse ordering instead of the perfect worst oracle; mean and standard
deviation are given for different groups of log files in
\tableref{Percentage increase in parsing time when using optimal ordering
instead of perfect best oracle table} and \tableref{Percentage decrease in
parsing time when using reverse ordering instead of perfect worst oracle
table}.

Once again, the difference between the oracles and optimal or reversed
orderings at its lowest when parsing log files resulting from a mail loop
(log files 22 \& 62--68), because the average number of rules tried per log
line is lower (see \graphref{Average number of rules tried per log line}).
The performance when parsing the first log file is different to the
remainder of the non-mail loop log files, because the hits field of every
rule is zero; optimal ordering is less efficient, and reverse ordering is
less inefficient.

Optimal ordering proves to be quite efficient: \tableref{Percentage
increase in parsing time when using optimal ordering instead of perfect
best oracle table} shows that optimal ordering is less than
\input{build/include-perfect-best-vs-optimal-mean} slower than parsing
using a magical oracle that divines the correct rule to use for each log
line.

\showgraph{build/graph-perfect-best-and-perfect-worst-and-optimal-and-reverse-vs-shuffled}{Parsing
time relative to shuffled ordering, including oracles}{Parsing time
relative to shuffled ordering including perfect best and worst graph}

\showtable{build/include-perfect-best-and-optimal-and-reverse-and-perfect-worst-vs-shuffle}{Parsing
time relative to shuffled ordering, including oracles}{Parsing time
relative to shuffled ordering, including oracles table}

\showgraph{build/graph-perfect-best-vs-optimal}{Percentage increase in
parsing time when using optimal ordering instead of perfect best
oracle}{Percentage increase in parsing time when using optimal ordering
instead of perfect best oracle graph}

\showtable{build/include-perfect-best-vs-optimal-stddev}{Percentage
increase in parsing time when using optimal ordering instead of perfect
best oracle}{Percentage increase in parsing time when using optimal
ordering instead of perfect best oracle table}

\showgraph{build/graph-perfect-worst-vs-reverse}{Percentage decrease in
parsing time when using reverse ordering instead of perfect worst
oracle}{Percentage decrease in parsing time when using reverse ordering
instead of perfect worst oracle graph}

\showtable{build/include-perfect-worst-vs-reverse-stddev}{Percentage
decrease in parsing time when using reverse ordering instead of perfect
worst oracle}{Percentage decrease in parsing time when using reverse
ordering instead of perfect worst oracle table}

@@

How any architecture scales as the number of rules increases is important,
but it is particularly important in this architecture because it is
expected that the typical parser will have a large number of rules.  There
are \numberOFrules{} rules in the full \parsername{} ruleset, whereas the
minimum number of rules required to parse the \numberOFlogFILES{} log files
used when generating the results in this thesis is \numberOFrulesMINIMUM{},
\numberOFrulesMINIMUMpercentage{} of the full ruleset.  The full ruleset is
larger because \parsername{} is tested with \numberOFlogFILESall{} log
files, to improve its parsing and ensure it correctly parses log files
generated by later Postfix versions.  A second set of test runs was
performed using the minimum ruleset, and the parsing times compared to
those generated using the full ruleset: the percentage parsing time
increase when using the full ruleset instead of the minimal ruleset for
optimal, shuffled and reverse orderings is shown in \graphref{Percentage
parsing time increase of maximum ruleset over minimum ruleset}.
